---
title: "myspotifymood"
author: "Alexa Pope"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
A package extending the capability of [spotifyr]
(https://github.com/charlie86/spotifyr) by measuring the valence (mood) and 
popularity of the tracks on the user's specified playlist. This README is 
derived from Matt Kearney's excellent 
[rtweet](https://github.com/mkearney/rtweet) 
documentation.

## Install

Install from GitHub with the following code:

```{r install, eval = FALSE}
if (!requireNamespace("devtools", quietly = TRUE)) {
  install.packages("devtools")
}
devtools::install_github("alexapope/myspotifymood")
```

First, set up a Dev account with Spotify to access their Web API here.
This will give you your Client ID and Client Secret. Once you have those, 
you can pull your access token into R with get_spotify_access_token().

The easiest way to authenticate is to set your credentials to the System 
Environment variables SPOTIFY_CLIENT_ID and SPOTIFY_CLIENT_SECRET. The 
default arguments to get_spotify_access_token() (and all other functions 
in this package) will refer to those. Alternatively, you can set them manually 
and make sure to explicitly refer to your access token in each subsequent 
function call.This authentication process is directly derived from Charlie 
Thompson's exceptionally detailed [spotifyr] 
(https://github.com/charlie86/spotifyr) authorization documentation. 

## Usage

This is the function that is currently live for myspotifymood. 

To begin, the user must first enter the following code, inserting their keys 
where appropriate:

```{setup instructions, eval = FALSE}
my_spotify_mood = function(username = "", client_id = "", 
                           client_secret = "", playlists = ""){ 
  Sys.setenv(client_id)
  Sys.setenv(client_secret)
  access_token = get_spotify_access_token()
```

Currently, this must be done at the start of every session.

Next, the fun begins with <code>myspotifymood</code>.

Its first argument takes the Spotify user's username and utilizes it to extract 
the user's necessary data.

The second and third argument take the user's Spotify authentication codes from 
Spotify API Developer (Client ID and Client Secret) in order for the user to be 
able to utilize SpotifyR.One would have to make link their Spotify account 
with Spotify Developer in order to retrieve these codes. 

The fourth argument determines the playlist id of what playlist the user will be
extracting data like tracks and track audio features from. This will output a 
data frame of all of the user's playlists and the user must locate their chosen
playlist's id as one of the variables in the data frame and input the code in 
lieu of the playlist argument. 

```{r usage, eval = FALSE}
## load my_spotify_mood
library(my_spotify_mood)

## Enter argument will be surrounded by quotation marks
my_spotify_mood(username, client_id, client_secret, playlists)

## In order to retrieve the playlist ID for the function 
playlists = spotifyr::get_user_playlists(username)

## Function results in a visualization of one's playlist track's valence.

```

Important potential variables in the playlists object are "name" the playlist 
name, "id" the playlist id, and "uri" different iteration of the playlist id. 
The uri is not necessary for this function so disregard it. 

The visualization, when the function is utilized, produces a column plot that 
all the tracks on the extracted playlist and the valence scores of the tracks 
with intervals of 0.25. The "x" legend to the right of the plot illustrates the 
degree of valance based on the values accompanied by lighter to darker shades of 
blue. The lighter and higher the color and valence score is, the happier the
is track is and viceversa for the darker and lower valence scores.

This function wishes to help it's user get a better understanding of their 
own music taste in accordance with the already established package by 
Charlie Thompson, [spotifyr] (https://github.com/charlie86/spotifyr). This
function cannot be utilized if the user doesn't have a Spotify account, and 
doesn't go through the authorization process. 


